Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-11 mr-review.Rmd
minor rewording of a sentence

Researchers have recognized two types of horizontal pleiotropy [@yuan_likelihood_2021]. The first occurs via 
exposure-independent paths. The resulting horizontal pleiotropic effects are independent of the
SNP-exposure relationships. The second type of horizontal pleiotropy manifests when 
in the presence of unobserved exposure-outcome confounding. It induces correlation between 
horizontal pleiotropic effects and SNP-exposure effects. Both types of horizontal pleiotropy 
violate standard MR modeling assumptions and can bias causal effect estimates and can increase false 
discoveries [@yuan_likelihood_2021].
Early MR analyses avoided confounding from horizontal pleiotropy by discarding 
instrumental SNPs that might be associated with the outcome 
[@verbanck_detection_2018;@zhu_causal_2018;@zeng_causal_2019-1;@zeng_causal_2019-2].
More recent methods have attempted to model horizontal pleiotropy 
[@morrison_mendelian_2020;@yuan_testing_2020;@yuan_likelihood_2021]. CAUSE [@morrison_mendelian_2020] and 
MRMix [@qi_mendelian_2019] both use a mixture of normal distributions to control for both 
types of horizontal pleiotropy. Modeling both types of horizontal pleiotropy is particularly 
challenging because the MR model likelihood often involves an integral that can't be solved analytically. 
Because of this issue with the model likelihood, both MRMix and CAUSE use other, non-likelihood-based methods for inference.
MRMix uses ...
CAUSE compares the out-of-sample prediction accuracies for two competing models that differ in inclusion of 
the causal effect parameter [@morrison_mendelian_2020]. Specifically, CAUSE calculates the expected log pointwise posterior density 
between the two models. 
MRAID models horizontal pleiotropy by ...
@yuan_likelihood_2021 found that MRMix is not robust to misspecification of SNP effect sizes and often is 
biased. CAUSE, they found, yields overly conservative p-values [@yuan_likelihood_2021].

>>>>

In the short time since publication of @verbanck_detection_2018, researchers have recognized two types of horizontal pleiotropy [@yuan_likelihood_2021]. The first occurs via 
exposure-independent paths. The resulting horizontal pleiotropic effects are independent of the
SNP-exposure relationships. The second type of horizontal pleiotropy manifests when 
in the presence of unobserved exposure-outcome confounding. It induces correlation between 
horizontal pleiotropic effects and SNP-exposure effects. Both types of horizontal pleiotropy 
violate standard MR modeling assumptions and can bias causal effect estimates and can increase false 
discoveries [@yuan_likelihood_2021].
Early MR analyses avoided confounding from horizontal pleiotropy by discarding 
instrumental SNPs that might be associated with the outcome 
[@verbanck_detection_2018;@zhu_causal_2018;@zeng_causal_2019-1;@zeng_causal_2019-2].
More recent methods have attempted to model horizontal pleiotropy 
[@morrison_mendelian_2020;@yuan_testing_2020;@yuan_likelihood_2021]. CAUSE [@morrison_mendelian_2020] and 
MRMix [@qi_mendelian_2019] both use a mixture of normal distributions to control for both 
types of horizontal pleiotropy. Modeling both types of horizontal pleiotropy is particularly 
challenging because the MR model likelihood often involves an integral that can't be solved analytically. 
Because of this issue with the model likelihood, both MRMix and CAUSE use other, non-likelihood-based methods for inference.
MRMix uses ...
CAUSE compares the out-of-sample prediction accuracies for two competing models that differ in inclusion of 
the causal effect parameter [@morrison_mendelian_2020]. Specifically, CAUSE calculates the expected pointwise log posterior density 
between the two models. 
MRAID models horizontal pleiotropy by ...
@yuan_likelihood_2021 found that MRMix is not robust to misspecification of SNP effect sizes and often is 
biased. CAUSE, they found, yields overly conservative p-values [@yuan_likelihood_2021].
